\documentclass[11pt]{article}
\usepackage[margin=1.1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{tikz}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue,
  pdftitle={Exact two-colour Rado numbers of x1+ax2-x3=c},
  pdfauthor={Jan Sherzad Colet},
  pdfkeywords={Rado numbers, Ramsey theory on the integers, monochromatic
    solutions, SAT solving, DRAT certificates}]{hyperref}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{remark}
\newtheorem{remark}{Remark}
\theoremstyle{definition}
\newtheorem{example}[theorem]{Example}

\newcommand{\Rad}{\mathrm{Rad}_2}
\newcommand{\vtwo}{\nu_2}
\DeclareMathOperator{\lcm}{lcm}

\title{Exact two-colour Rado numbers of $x_1 + a x_2 - x_3 = c$: \\
the window $0 < c < a$ and two conjectures \\ of Dwivedi and Tripathi}
\author{Jan Sherzad Colet\thanks{Independent researcher.
\texttt{jansherzadcolet@gmail.com}. Every computational
claim in this note is accompanied by machine-checkable certificates (witness
colourings and DRAT proofs verified by \texttt{drat-trim}); the complete
verification package is available from the author and at
\url{https://github.com/4Gatitos/RadoNumbers}.}}
\date{July 2026}

\begin{document}
\maketitle

\begin{abstract}
For integers $a \ge 1$ and $c$, the $2$-colour Rado number $\Rad(a;c)$ is the
least $N$ such that every $2$-colouring of $\{1,\dots,N\}$ admits a
monochromatic solution of $x_1 + a x_2 - x_3 = c$, repeated values being
allowed; it exists if and only if $\vtwo(a) \le \vtwo(c)$, where $\vtwo$ is
the $2$-adic valuation. Dwivedi and Tripathi
[\emph{Integers} \textbf{20} (2020), \#A36] determined $\Rad(a;c)$ in several
parameter ranges and asked for the remaining exact values. We resolve the
window $0 < c < a$ completely: writing $d = a-c$ and $r \equiv a \pmod{2d}$
with $r \in [1,2d]$,
\[
\Rad(a;\,a-d) \;=\;
\begin{cases}
2a+2, & d = 1,\\[2pt]
(a+3)d + a - \min(r-1,\, d-1), & d \ge 2,
\end{cases}
\]
whenever the number exists. The upper-bound argument extends to $c \le 0$,
proving Conjecture~1 of Dwivedi and Tripathi in full; their Conjecture~2
then follows by a reflection argument. In every regime the final obstruction
is precisely the $2$-adic existence condition. In the remaining band
$a < c < a(K+2)$, where
$K+1 = \lceil (1+c(a+3))/(1+a(a+3)) \rceil$ is the general lower bound of
Dwivedi and Tripathi, the values appear to follow no closed formula; there
we prove an extension of their multiples theorem and announce the complete
determination of the band, with proofs to appear in a companion manuscript.
All results are corroborated by more than $30{,}000$ SAT-computed values
with independently verifiable certificates.

\medskip
\noindent\emph{2020 Mathematics Subject Classification:} 05D10 (primary);
11B75, 68R05 (secondary).

\noindent\emph{Keywords:} Rado numbers, Ramsey theory on the integers,
monochromatic solutions, SAT solving, DRAT certificates.
\end{abstract}

\section{Introduction}

For an equation $\mathcal{E}$ in the variables $x_1,\dots,x_m$, the
$2$-colour Rado number $\Rad(\mathcal{E})$ is the least positive integer $N$
such that every $2$-colouring of the integer interval $[1,N]$ contains a
monochromatic solution of $\mathcal{E}$, that is, a solution all of whose
entries receive the same colour; solutions with repeated values are allowed.
Following Dwivedi and Tripathi \cite{DT2020}, we study the family
\[
\mathcal{E}_{a,c}\colon\quad x_1 + a x_2 - x_3 = c ,
\]
where $a$ is a positive integer and $c$ any integer, and write
$\Rad(a;c) := \Rad(\mathcal{E}_{a,c})$.

Dwivedi and Tripathi proved that $\Rad(a;c)$ exists if and only if
$\vtwo(a) \le \vtwo(c)$ (with $\vtwo(0) = \infty$), and determined it exactly
in several parameter ranges: for $c \le 0$ when $a \mid c$, or when $a$ is odd
and $c$ is sufficiently negative (in a precise residue-dependent sense); for
$c = ma$ with $1 < m \le a+1$; and for $c > a$ sufficiently large. In the
remaining ranges they gave upper and lower bounds, together with two
conjectures for the zones adjacent to their theorems. In the recent paper
\cite{DT2025} the same authors returned to the problem and wrote that
``it would be desirable to determine the exact value of the Rado number in
all cases.''

In the window $0 < c < a$ their results give the bounds
$(a+3)(a-c)+1 \le \Rad(a;c) \le (2a+1)(a-c)+1$ (for odd $a$; similarly for
even $a$), and no conjecture. A subsequent paper of Arora, Dwivedi and
Tripathi \cite{ADT2025} extends the lower bound to a general class of
equations $\sum_i a_i x_i - x_m = c$, with matching exact values when the
coefficient multiset is \emph{$2$-distributable}; for the equation studied
here the multiset is $\{1, a\}$, which is $2$-distributable only for
$a \le 2$, so in our window that paper again yields exactly the lower bound
$(a+3)(a-c)+1$. This note closes the window completely --- in particular,
the lower bound of \cite{DT2020,ADT2025} is never tight there for
$a \ge 3$, $d \ge 2$.

\begin{theorem}\label{thm:d1}
Let $a \ge 3$ be odd. Then $\Rad(a;\,a-1) = 2a+2$.
\end{theorem}

(For even $a$ the number $\Rad(a;a-1)$ does not exist, since $a-1$ is odd;
so Theorem~\ref{thm:d1} covers the whole diagonal $c = a-1$. The case $a=3$,
i.e.\ $\Rad(3;2) = 8$, appears in \cite{DT2020}; the general statement is
new.)

\begin{theorem}\label{thm:main}
Let $a > d \ge 2$ with $\vtwo(a) \le \vtwo(a-d)$, and write $a = 2dq + r$
with $q \ge 0$ an integer and $r \in [1, 2d]$. Then
\[
\Rad(a;\,a-d) \;=\; (a+3)d + a - \min(r-1,\, d-1).
\]
\end{theorem}

Under the existence condition one automatically has $r \in [1, 2d-1]$
(Lemma~\ref{lem:norm}), and the two theorems together determine
$\Rad(a;c)$ for every pair with $0 < c < a$ for which it exists. It is
curious that the uniform formula of Theorem~\ref{thm:main} fails for $d = 1$
by exactly one unit: the extremal colouring family of
Section~\ref{sec:family} degenerates at $d=1$ into a colouring of
$[1, 2a+2]$ which is \emph{not} valid --- and the proof of
Theorem~\ref{thm:d1} shows that nothing else works at that length either.

Our third and fourth results settle the two conjectures of \cite{DT2020}.
For $c \le 0$ one has $d = a - c \ge a$, so that $r \equiv a \pmod{2d}$ is
simply $r = a$: the window formula of Theorem~\ref{thm:main} formally
reduces to $(a+3)d + a - (a-1) = (a+3)(a-c) + 1$, which is exactly the value
predicted by Conjecture~1 of \cite{DT2020}. This is no coincidence: our
upper-bound proof for the regime $r \le d$ extends to this degenerate case
($q = 0$ in the normalisation $a = 2dq + r$) after adjusting two range
conditions, and the lower bound is Theorem~2 of \cite{DT2020}.

\medskip
\noindent\textbf{Theorem~\ref{thm:c1}} (Conjecture~1 of \cite{DT2020}).
\emph{Let $a \ge 1$ and $c \le 0$ with $\vtwo(a) \le \vtwo(c)$. Then
$\Rad(a;c) = (a+3)(a-c) + 1$.}

\medskip
\noindent\textbf{Theorem~\ref{thm:c2}} (Conjecture~2 of \cite{DT2020}).
\emph{Let $a \ge 1$ and let $c \ge a(K+2)$ with $\vtwo(a) \le \vtwo(c)$, where
$K = \bigl\lceil (1 + c(a+3)) / (1 + a(a+3)) \bigr\rceil - 1$. Then
$\Rad(a;c) = K + 1$.}

\medskip
Theorem~\ref{thm:c2} follows from Theorem~\ref{thm:c1} by a
reflection argument generalising the proof of Theorem~7 of \cite{DT2020}
(Section~\ref{sec:conj}). Taken together, these results determine
$\Rad(a;c)$ exactly for every $c \le 0$, every $0 < c \le a$, and every
$c \ge a(K+2)$. The remaining zone --- the band $a < c < a(K+2)$ --- behaves
very differently: no closed formula holds there, the fluctuations being of
Sturmian type. Section~\ref{sec:band} states what we can prove about the
band: an extension of the multiples theorem of \cite{DT2020}
(Theorem~\ref{thm:band-mult}, proved here), the complete determination of
the band's deep tail, where the threshold is itself a Rado number from the
window $0 < c < a$ (Theorem~\ref{thm:band-tail}), and the exact
determination of the whole band by a polynomial-time algorithm
(Theorem~\ref{thm:band-alg}); the proofs of the last two will appear in the
companion manuscript \cite{companion}.

Table~\ref{tab:summary} and Figure~\ref{fig:zones} summarise the resulting
picture.

\begin{table}[ht]
\centering
\begin{tabular}{lll}
\hline
zone & $\Rad(a;c)$ & reference \\
\hline
$c \le 0$ & $(a+3)(a-c)+1$ & Theorem~\ref{thm:c1} \\
$0 < c < a$, $d = 1$ & $2a+2$ & Theorem~\ref{thm:d1} \\
$0 < c < a$, $d \ge 2$ & $(a+3)d + a - \min(r-1,d-1)$ & Theorem~\ref{thm:main} \\
$c = a$ & $1$ & \cite{DT2020} \\
$a < c < a(K+2)$, $a \mid c$ & $c/a$ & Theorem~\ref{thm:band-mult} \\
$a < c < a(K+2)$, $a \nmid c$ & deep-tail formula; exact algorithm &
Theorems~\ref{thm:band-tail}--\ref{thm:band-alg}$^{*}$ \\
$c \ge a(K+2)$ & $K+1$ & Theorem~\ref{thm:c2} \\
\hline
\end{tabular}
\caption{The value of $\Rad(a;c)$ on each parameter zone, assuming the
existence condition $\vtwo(a) \le \vtwo(c)$. Here $d = a-c$,
$r \equiv a \pmod{2d}$ with $r \in [1,2d]$, and
$K = \lceil (1+c(a+3))/(1+a(a+3)) \rceil - 1$. Entries marked $^{*}$ are
announced here and proved in the companion manuscript; all other entries
are proved in this note or in \cite{DT2020}.}
\label{tab:summary}
\end{table}

\begin{figure}[ht]
\centering
\begin{tikzpicture}[xscale=0.92, yscale=0.9, font=\small]
% axis
\draw[->] (-3.4,0) -- (12.6,0) node[right] {$c$};
% zone boundaries
\foreach \x/\lab in {0/{$0$}, 2.2/{$a$}, 8.2/{$a(K{+}2)$}}
  \draw (\x,-0.12) -- (\x,0.12) node[below=8pt] {\lab};
% zone bands
\fill[gray!14] (-3.3,0.25) rectangle (0,1.05);
\fill[gray!28] (0,0.25) rectangle (2.2,1.05);
\fill[gray!45] (2.2,0.25) rectangle (8.2,1.05);
\fill[gray!14] (8.2,0.25) rectangle (12.5,1.05);
% zone labels (formulas)
\node at (-1.65,0.65) {$(a{+}3)(a{-}c){+}1$};
\node at (1.1,0.65) {window};
\node at (5.2,1.35) {middle band: Sturmian fluctuations above $K{+}1$};
\node at (10.35,0.65) {$K+1$};
% theorem tags
\node at (-1.65,-0.75) {Thm~\ref{thm:c1}};
\node at (1.1,-0.75) {Thms~\ref{thm:d1}--\ref{thm:main}};
\node at (5.2,-0.75) {Thms~\ref{thm:band-mult}--\ref{thm:band-alg}};
\node at (10.35,-0.75) {Thm~\ref{thm:c2}};
% sawtooth inside the band, above the K+1 baseline
\draw[thick] (2.3,0.42) -- (2.8,0.95) -- (3.0,0.55) -- (3.6,0.9) -- (3.8,0.5)
  -- (4.5,0.85) -- (4.7,0.47) -- (5.5,0.8) -- (5.7,0.44) -- (6.6,0.72)
  -- (6.8,0.4) -- (7.6,0.6) -- (7.8,0.36) -- (8.15,0.33);
\draw[dashed] (2.2,0.3) -- (8.2,0.3);
\end{tikzpicture}
\caption{The zones of $\Rad(a;c)$ (schematic, $c$ increasing to the right).
Outside the middle band the value follows the closed formulas of
Table~\ref{tab:summary}; inside it, the value oscillates above the lower
bound $K+1$ of \cite{DT2020} (dashed) in a quasi-periodic, three-distance
pattern, and is determined by the algorithm of Theorem~\ref{thm:band-alg}.}
\label{fig:zones}
\end{figure}

The proofs have a common architecture, which generalises the forcing
arguments of \cite{DT2020}. The lower bounds are witnessed by an explicit
two-parameter family of periodic colourings with regularly spaced phase
defects (Section~\ref{sec:family}). The upper bounds are rigidity arguments:
any putative valid colouring of the critical interval is forced, triple by
triple, into a highly structured pattern, and the final obstruction that
kills the pattern is, in both parameter regimes, \emph{exactly} the $2$-adic
existence condition $\vtwo(a) \le \vtwo(a-d)$ --- in the regime $r \le d$
through the fact that $\gcd(r, 2d) \mid d$, and in the regime $r > d$
through the incompatibility of antipodal and $s$-antiperiodic patterns on
$\mathbb{Z}_{2d}$ when $\vtwo(s) \ne \vtwo(d)$, $s = r-d$
(Sections~\ref{sec:ub1} and~\ref{sec:ub2}).

\paragraph{Methodology and verifiability.}
The formula of Theorem~\ref{thm:main}, the extremal family, and the key
structural lemmas were first \emph{discovered} through an extensive
SAT-solver-based computation of exact Rado numbers: $3662$ pairs $(a,c)$
with $3 \le a \le 15$, $-100 \le c \le 300$, later extended along $0 < c < a$
up to $a = 45$. Every computed value $\Rad(a;c) = N$ is certified in both
directions: a valid colouring of $[1,N-1]$ (checkable by a trivial
independent program) and a DRAT proof of unsatisfiability at $N$, generated
by CaDiCaL~2.1.3 \cite{cadical} and verified by \texttt{drat-trim} against an independently
regenerated CNF encoding. The proofs presented here are, however, entirely
self-contained and human-readable; the computations play no logical role in
the theorems of this note: they served to discover the statements and to
corroborate them independently. Every one of the $925$ computed values with
$c \le 0$, and every value in the ranges of the four theorems, matches the
formulas proved here (Section~\ref{sec:comp}).

\section{Preliminaries}\label{sec:prelim}

Throughout, $[u,v]$ denotes $\{u, u+1, \dots, v\}$ and $\vtwo$ the $2$-adic
valuation. A \emph{colouring} is a map $\chi\colon [1,N] \to \{0,1\}$; it is
\emph{valid} (for the pair $(a,c)$) if no solution
$(x_1,x_2,x_3) \in [1,N]^3$ of $x_1 + a x_2 - x_3 = c$ is monochromatic.

Solving for $x_3$, we see that the solutions in $[1,N]^3$ are exactly the
triples
\begin{equation}\label{eq:triples}
T(x,k) \;=\; \bigl(x,\; k,\; x + a(k-1) + d\bigr), \qquad d := a - c,
\end{equation}
with $x, k \in [1,N]$ and $1 \le x + a(k-1) + d \le N$ (for $d \ge 1$, which
holds everywhere outside Section~\ref{sec:band}, the constraint
$x + a(k-1) + d \ge 1$ is automatic). We record two standard facts.

\begin{lemma}[monotonicity]\label{lem:mono}
If $[1,N]$ admits a valid colouring, so does $[1,N']$ for every $N' \le N$
(restrict the colouring). Hence $\Rad(a;c) > N$ if and only if $[1,N]$
admits a valid colouring, and to determine $\Rad(a;c) = N$ it suffices to
exhibit a valid colouring of $[1,N-1]$ and to prove that no valid colouring
of $[1,N]$ exists.
\end{lemma}

\begin{lemma}[colour symmetry]\label{lem:wlog}
$\chi$ is valid if and only if $1-\chi$ is valid. In any proof by
contradiction we may therefore normalise $\chi(1) = 0$.
\end{lemma}

\begin{lemma}[normalisation]\label{lem:norm}
Let $d \ge 1$, $a > d$, and write $a = 2dq + r$ with $r \in [1, 2d]$.
\begin{enumerate}
\item[(i)] If $r = 2d$, then $\vtwo(a) > \vtwo(a-d)$, so $\Rad(a;a-d)$ does
not exist. Under the existence condition, therefore, $r \in [1, 2d-1]$.
\item[(ii)] If $r \le d$ then $q \ge 1$ (so $a \ge 2d+1$).
\item[(iii)] If $r = d + s$ with $s \in [1, d-1]$, then $q \ge 0$, and the
existence condition is equivalent to $\vtwo(s) \ne \vtwo(d)$
(Lemma~\ref{lem:2adic-b} below); if $r \le d$, the existence condition
implies $\vtwo(r) \le \vtwo(d)$, hence $\gcd(r,2d) \mid d$
(Lemma~\ref{lem:2adic-a} below).
\end{enumerate}
\end{lemma}

\begin{proof}
(i) Write $a = 2dm$, $m \ge 1$. Then $\vtwo(a) \ge \vtwo(d)+1$ while
$a - d = d(2m-1)$ gives $\vtwo(a-d) = \vtwo(d)$.
(ii) If $q = 0$ then $a = r \le d$, contradicting $a > d$.
(iii) The two $2$-adic statements are proved as
Lemmas~\ref{lem:2adic-a} and~\ref{lem:2adic-b} in the sections where they
are used.
\end{proof}

\section{The diagonal $c = a-1$: proof of Theorem~\ref{thm:d1}}\label{sec:d1}

Here $d = 1$ and the solution triples \eqref{eq:triples} are
$T(x,k) = (x, k, x + a(k-1) + 1)$.

\subsection*{Lower bound}
Colour $[1, 2a+1]$ by the classes
\[
A \;=\; \{\,x \text{ odd} : x \le a\,\} \cup \{\,x \text{ even} : a+3 \le x \le 2a\,\},
\qquad
B \;=\; [1, 2a+1] \setminus A .
\]
If a triple $T(x,k)$ lies in $[1, 2a+1]^3$ then $a(k-1) \le 2a-1$, so
$k \in \{1,2\}$. A monochromatic triple in $A$ would need $k = 1$ (as
$2 \notin A$: the even elements of $A$ start at $a+3$); but for $x \in A$ odd,
$x+1$ is even and $\le a+1$, hence in $B$, and for $x \in A$ even,
$x+1$ is odd and $\ge a+4$, hence in $B$. A monochromatic triple in $B$ would
need $k = 2$ (as $1 \notin B$); but for $x \in B$ even, $x \le a+1$: if
$x \le a-1$ then $x_3 = x + a + 1$ is even in $[a+3, 2a]$, hence in $A$,
and if $x = a+1$ then $x_3 = 2a+2$ is out of range; for $x \in B$ odd,
$x \ge a+2$, so $x_3 \ge 2a+3$ is out of range. Hence the colouring is valid
and $\Rad(a; a-1) \ge 2a+2$.

\subsection*{Upper bound}
Let $\chi$ colour $[1, 2a+2]$ with no monochromatic triple; by
Lemma~\ref{lem:wlog}, $\chi(1) = 0$.

\emph{Step 1.} $T(x,1) = (x,1,x+1)$ is a solution for every $x \le 2a+1$.
Since $\chi(1) = 0$, no two consecutive integers are both coloured $0$.
In particular $T(1,1) = (1,1,2)$ forces $\chi(2) = 1$.

\emph{Step 2.} $T(x,2) = (x,2,x+a+1)$ is a solution for $x \le a+1$. Since
$\chi(2) = 1$: if $\chi(x) = 1$ with $x \le a+1$, then $\chi(x+a+1) = 0$.

\emph{Step 3.} The pairs $\{x, x+a+1\}$, $x \in [1, a+1]$, partition
$[1, 2a+2]$. By Step~2 each pair contains a $0$, so $\#\chi^{-1}(0) \ge a+1$;
by Step~1, $\#\chi^{-1}(0) \le a+1$. Hence exactly $a+1$ zeros, no two
adjacent, in an interval of length $2a+2$: the zero set must be
\[
Z_j \;=\; \{1, 3, \dots, 2j-3\} \cup \{2j, 2j+2, \dots, 2a+2\}
\]
for some $j$, and $\chi(1) = 0$ forces $j \in [2, a+2]$.

\emph{Step 4.} Each $j$ is contradictory.
If $j = a+2$ (zeros are all the odds), the triple $(a+1, 2, 2a+2)$ is
monochromatic of colour $1$ ($a+1$ and $2a+2$ are even).
If $3 \le j \le a+1$, then $\chi(1) = \chi(3) = 0$ and $2a+2 \ge 2j$ is an
even member of $Z_j$; the triple $(1, 3, 2a+2)$ is monochromatic of colour
$0$.
If $j = 2$ (zeros are $\{1\} \cup \{4, 6, \dots, 2a+2\}$), then
$\chi(3) = \chi(2) = 1$ and $a+4$ is odd and $\ge 5$, hence coloured $1$;
the triple $(3, 2, a+4)$ is monochromatic of colour $1$ (note
$a + 4 \le 2a+2$ for $a \ge 2$). \qed

\section{The extremal family and the lower bound for $d \ge 2$}\label{sec:family}
\input{sec_lb}

\section{The upper bound in the regime $r \le d$}\label{sec:ub1}
\input{sec_ub1}

\section{The upper bound in the regime $r > d$}\label{sec:ub2}
\input{sec_ub2}

\section{The conjectures of Dwivedi and Tripathi}\label{sec:conj}
\input{sec_conj}

\section{The middle band}\label{sec:band}

Theorems~\ref{thm:d1}, \ref{thm:main}, \ref{thm:c1} and~\ref{thm:c2}
determine $\Rad(a;c)$ everywhere
except in the \emph{middle band} $a < c < a(K+2)$ (with $K$ as in
Theorem~\ref{thm:c2}), where the behaviour is qualitatively different from
that in all the other zones. About the band, the
status of the statements below is as follows.
\emph{Theorem~\ref{thm:band-mult} is proved here in full.}
\emph{Theorems~\ref{thm:band-tail} and~\ref{thm:band-alg} are announced}:
we state them precisely and describe the structure of their proofs, but the
complete arguments are longer and will appear in the companion manuscript
\cite{companion}; all supporting computations and verification programs are
already available with this note. Nothing in
Sections~\ref{sec:prelim}--\ref{sec:conj} depends on
Theorems~\ref{thm:band-tail} or~\ref{thm:band-alg}.

The multiples of $a$ inside the band are governed by a simple degeneracy,
which extends Theorem~5 of \cite{DT2020} (proved there for $1 < m \le a+1$)
to the full range in which the value $m$ is attained.

\begin{theorem}\label{thm:band-mult}
Let $a \ge 1$ and let $c = ma$ with $2 \le m \le a(a+3)+1$. Then
$\Rad(a;\,ma) = m$.
\end{theorem}

\begin{proof}
The triple $(x_1,x_2,x_3) = (m,m,m)$ satisfies $m + am - m = am = c$ and is
monochromatic under every colouring, so $\Rad(a;ma) \le m$. For the lower
bound, Theorem~6 of \cite{DT2020} gives $\Rad(a;c) \ge K+1$, and with
$M := a(a+3)+1$ one computes $1 + ma(a+3) = mM - (m-1)$, whence
$K + 1 = \lceil (mM - (m-1))/M \rceil = m$ for $m \le M$.
\end{proof}

For $c$ deep inside the band the value is pinned to the immediate
neighbourhood of the lower bound of \cite{DT2020}, and the exact threshold
turns out to be --- remarkably --- a Rado number from the window
$0 < c < a$, so that the band inherits its fine structure from
Theorems~\ref{thm:d1} and~\ref{thm:main}.

\begin{theorem}[deep tail]\label{thm:band-tail}
Let $(a,c)$ lie in the middle band with $\vtwo(a) \le \vtwo(c)$ and
$a \nmid c$, write $t = c \bmod a \in [1, a-1]$ and
$\lambda = \lceil c/a \rceil$, and set
\[
E_1 = E_1(a,t) \;:=\; \Rad(a;\,a-t),
\]
the value determined by Theorem~\ref{thm:d1} (for $t = 1$) and
Theorem~\ref{thm:main} (for $t \ge 2$); this Rado number exists, since
$\vtwo(a) \le \vtwo(c)$ forces $\vtwo(a) \le \vtwo(t)$, hence
$\vtwo(a) \le \vtwo(a-t)$. If $t \ge 2$ and
$\lambda \ge t(a+3)+2$, or $t = 1$ and $\lambda \ge 2a-2$, then
\[
\Rad(a;c) \;=\; \lambda - 1 + [\lambda \le E_1],
\]
where $[P]$ equals $1$ if the condition $P$ holds and $0$ otherwise.
\end{theorem}

In the remaining (shallow) part of the band no closed formula appears to
hold, and the computations indicate that this is not an artefact of our
methods. Along each diagonal
$\{c : c \equiv -\mu \pmod a\}$ the excess $\Rad(a;c) - (K+1)$ forms a
non-increasing staircase whose step pattern follows the three-distance
(Sturmian) structure of the rotation by $c - a$ on $\mathbb{Z}_a$, with
step sizes governed by the continued-fraction expansion of $\mu/a$; the
staircases are eventually periodic only for special $\mu$. A concrete
finite obstruction: for $(a,c) = (7,16)$ \emph{every} valid colouring of
the extremal interval $[1, 15]$ has at least nine maximal runs (an
exhaustive check), so no family of colourings with a bounded number of
blocks can witness the lower bounds uniformly across the band. The exact
values can nonetheless be determined:

\begin{theorem}[exact determination of the band]\label{thm:band-alg}
There is an algorithm which, given $(a,c)$ in the middle band with
$\vtwo(a) \le \vtwo(c)$, computes $\Rad(a;c)$ exactly in time polynomial in
$a$.
\end{theorem}

The algorithm decides, for each candidate length $N$, the existence of a
valid colouring of $[1,N]$ as follows. For $a \nmid c$ (the multiples of
$a$ being settled by Theorem~\ref{thm:band-mult}), all solution triples of
\eqref{eq:triples} share the multiplier window
$W = \{k \in [1,N] : 1 \le |ak - c| \le N-1\}$, a set of $O(N/a)$
consecutive integers; once the colours of $W$ are fixed, validity is a
$2$-{\upshape SAT} instance; and it suffices to test the $O(|W|^2)$
assignments in which one colour class of $W$ is a single contiguous block.
The completeness of this restricted family --- every colourable length
admits a valid colouring of the restricted form --- is the main technical
result behind the theorem: in any valid colouring the two central
multipliers of $W$ must receive the same colour, and every non-extremal
multiplier inherits that colour, by a rigidity argument whose only
arithmetic input is, once again, the existence condition
$\vtwo(a) \le \vtwo(c)$.

Across the $28{,}000$-plus band values computed to date the algorithm
agrees with the SAT-computed ground truth without exception, and each value
carries the same two-sided certificates as the rest of this note.

\begin{example}\label{ex:716}
Take $(a,c) = (7,16)$, so $t = 2$, $\lambda = 3$, $\mu = 5$ and $K+1 = 3$,
while in fact $\Rad(7;16) = 16$ (a certified value; see
Section~\ref{sec:comp}): the lower bound of \cite{DT2020} is off by
$13$ here. An extremal colouring of $[1,15]$ is
\[
1\;0\;0\;1\;1\;0\;1\;1\;1\;0\;1\;1\;0\;0\;1,
\]
which has nine maximal runs --- and, as noted above, no valid colouring of
$[1,15]$ has fewer. At $N = 15$ the multiplier window is
$W = \{1,2,3,4\}$, with shifts $7k - 16$ equal to $-9, -2, 5, 12$; the two
central multipliers $k = 2, 3$ carry the smallest distances $t = 2$ and
$\mu = 5$. The restriction of the colouring above to $W$ is $(1,0,0,1)$ ---
just three runs, exactly the shape required by the restricted family of
Theorem~\ref{thm:band-alg}. The colouring is globally complex, but its
window trace is simple; this contrast is the heart of the band.
\end{example}

\section{Computations and certificates}\label{sec:comp}

The exact values of $\Rad(a;c)$ were computed for all existing pairs with
$3 \le a \le 15$ and $-100 \le c \le 300$ ($3662$ pairs), and additionally
for all existing pairs with $0 < c < a$ up to $a = 45$. The encoding is the
obvious one: a Boolean variable per integer in $[1,N]$ and, for every
solution triple \eqref{eq:triples}, two clauses forbidding the two
monochromatic colourings of that triple. For each pair, the exact value
$\Rad(a;c) = N$ is certified by:
\begin{enumerate}
\item a valid colouring of $[1, N-1]$, re-verified by an independent
brute-force checker of about $30$ lines that has no code in common with the
search program; and
\item a DRAT proof of unsatisfiability of the instance at $[1,N]$, produced
by CaDiCaL~2.1.3 and verified with \texttt{drat-trim}
\cite{drattrim} against a CNF regenerated by a second, independently written
encoder.
\end{enumerate}
As validation, the same programs reproduce all $42$ values proved in
\cite{DT2020} that fall in the computed range, as well as every value in the
ranges covered by Theorems~\ref{thm:d1} and~\ref{thm:main}.

Beyond the window $0 < c < a$, the computations corroborate
Theorems~\ref{thm:c1} and~\ref{thm:c2} at every computed point:
all $925$ values with $c \le 0$ in the grid (including the $234$ cases that
were open before this note) match $(a+3)(a-c)+1$, as do $8$ further
out-of-sample values with $a$ up to $33$ computed after the fact; and all
$300$ computed values with $c \ge a(K+2)$ match $K+1$ ($12$ of them lie
outside the ranges covered by the partial results of \cite{DT2020}).

In the middle band the computations yield over $28{,}000$ further exact
values (including complete sweeps of the band for several values of $a$),
against which the results of Section~\ref{sec:band} were checked without
exception.

All witnesses, DRAT certificates, the two independent encoders, the checker,
and scripts reproducing every finite verification used in the proofs are
available from the author and in the repository accompanying this note,
\url{https://github.com/4Gatitos/RadoNumbers}.

\begin{thebibliography}{9}

\bibitem{DT2020}
S.~Dwivedi and A.~Tripathi,
On the two-color Rado number for $x_1 + a x_2 - x_3 = c$,
\emph{Integers} \textbf{20} (2020), \#A36.
\url{http://math.colgate.edu/~integers/u36/u36.pdf}

\bibitem{DT2025}
S.~Dwivedi and A.~Tripathi,
On the two-color disjunctive Rado number for the equations
$\sum_{i=1}^{m-2} x_i + a x_{m-1} - x_m = c_j$,
\emph{Integers} \textbf{25} (2025), \#A108.
\url{http://math.colgate.edu/~integers/z108/z108.pdf}

\bibitem{ADT2025}
N.~Arora, S.~Dwivedi and A.~Tripathi,
Improved two-colour Rado numbers for linear equations with certain
coefficients,
\emph{Electron.\ J.\ Combin.} \textbf{32}(4) (2025), \#P4.63.
\url{https://doi.org/10.37236/13596}

\bibitem{cadical}
A.~Biere, T.~Faller, K.~Fazekas, M.~Fleury, N.~Froleyks and F.~Pollitt,
CaDiCaL 2.0,
in \emph{Computer Aided Verification (CAV 2024)}, LNCS 14681, Springer,
2024, 133--152.

\bibitem{companion}
J.~Sherzad Colet,
The middle band of the two-colour Rado numbers of $x_1 + a x_2 - x_3 = c$,
in preparation.

\bibitem{drattrim}
N.~Wetzler, M.~J.~H. Heule and W.~A. Hunt~Jr.,
DRAT-trim: efficient checking and trimming using expressive clausal proofs,
in \emph{SAT 2014}, LNCS 8561, Springer, 2014, 422--429.
\url{https://doi.org/10.1007/978-3-319-09284-3_31}

\end{thebibliography}

\end{document}
